The Impact of Model Assumptions on RES Drought Analysis

### 2024-11-15 Article/main.tex
Update Article

\textcolor{magenta}{Some things to keep in mind when reading from Damian's comments:
	1. The whole current/projected and existing/balanced thing. Pick one and stick to it, including the figures (also fix the a/c level in fig 10)
	2. See if you are happy with the section titles throughout the document
	3. Check if when it is EirGrid/SONI it is clear, both in the text and references
	4. Is the Ireland / ROI / NI distinction clear enough? I did not see the line where we say that we call ROI+NI=Ireland (may have missed it) but I am alright with how it is written now either way}

@@

This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios. The first represents the current energy mix, predominantly driven by wind, while the second is based on projected energy scenarios, featuring a more balanced share between wind and PV. A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.

@@

This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy. The primary data sources include EirGrid, the transmission system operator for the Republic of Ireland, and the ERA5 reanalysis dataset.

@@

For the second dataset, we use the gridded dataset from C3S-E to create a CF time series which is representative of the actual RES farm distribution. Using the farms' coordinates, we identify the nearest grid point on the C3S-E grid for each farm. We retrieve the CF values from the C3S-E dataset for its corresponding grid point. We then average the CF for each farm using the installed capacity as weights. This process results in a CF time series for each renewable energy source (wind energy and PV energy), which is representative of the location of the RES farms in Ireland.

@@

This study investigates wind and PV capacity factors (CF) using three different models, resulting in six individual CF time series based on the installed capacities of 2023. In addition to analysing wind and PV generation separately, a combined CF is computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV). This results in a total of nine CF time series for the existing scenario, which represents the current distribution of RES installations in Ireland.

Given that PV capacity in Ireland remains relatively low in 2023, and in order to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario, where the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW. These figures are based on the target energy capacity outlined in the roadmap published by EirGrid~\cite{eirgrid2023future}. In this scenario, new PV time series are generated for both the Atlite and C3S-E G models, while the wind CF time series remain the same, as significant changes in wind farm distribution are not expected. From there, the results of these new time series are combined using updated weights which reflect a more balanced scenario (9 GW for wind and 8 GW for PV). \textcolor{magenta}{Is the specific distribution of the capacities mentioned somewhere? Same locations but scaled for wind, change in locations for PV}

@@

\begin{figure}[ht!]
	\centering
	\includegraphics[width=\textwidth]{droughts_methodology}
	\caption{Wind time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021. The black dashed line indicates the CF threshold. The grey bar shows the period identified as a wind drought under our definition}
	\label{fig:find_res_droughts}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{power_curve}
	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
	\label{fig:power_curve}
\end{figure}

The smoothing of the wind turbine power curve represents losses due to the wake effect \textcolor{magenta}{yes, but not only. we do not want someone to be able to suck us into what the smoothing means rabbithole. My proposal: The smoothing of the wind turbine power curve represents losses associated with each individual turbine and the interaction between the different turbines}, which is important when modelling wind energy on larger scales. The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.

@@

\textcolor{magenta}{I was reading Damian's comment about the tables being "a lot of numbers". I think the way it is written now is clear but please review it and see if you are happy with it too. I though about maybe explaining that RMSE and MBE are better closer to 0 and CC closer to 1, but honestly I think that the reader will know that, it is quite a common method and not a complex selection of skill scores.}

@@

The median, first, and third quartiles for the Atlite dataset are consistently greater than or equal to those of the other two datasets, regardless of the duration range or type of renewable energy considered. This difference arises from the wind turbine power curve model used in the C3S-E datasets, which tends to overestimate the wind CF (Fig.~\ref{fig:wind_verification_contour}). As a result, the overall number of RES droughts is underestimated in the C3S-E datasets compared to Atlite. \textcolor{magenta}{Sorry to add this here again, but where are we on the current/projected/balanced thing? Just please make sure it is consistent, we can figure it out once Conor has gone over it, it's not crucial to the reading but I just remembered.}

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_number_events}
	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events. The boxes display the first and third quartiles and the median is marked by a black line. The whiskers indicate the 5th and 95th percentiles}
	\label{fig:boxplot_number_events}	
\end{figure}

@@

For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves significantly different than the two C3S-E datasets. The Atlite results follow the ones for wind with a log-linear increase, but reach higher values in general with the longest event lasting forty days. For C3S-E G and C3S-E N, the duration of RES droughts increases in a log-linear pattern for events lasting less than 16 days. Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period. This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months. The difference between Atlite and the C3S-E datasets likely arises from how each model handles PV output near the threshold. Atlite appears to hover slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events. In contrast, C3S-E G and C3S-E N tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter. This sensitivity to the threshold highlights how slight model differences can have substantial effects on drought duration estimates, particularly for PV in low-generation conditions. \textcolor{magenta}{Not sure if it is worth a mention, but in case you had not made the connection: this is consistent with atlite having the highest MBE out of the three in the verification}

For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), reinforcing that wind is the primary driver of drought patterns in the current energy mix. In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations. For example, the return period for a five-day drought event extends from roughly six months to four years in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets. The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one. This convergence is due to the reduced proportion of wind capacity in the balanced scenario, which lessens the impact of wind-specific characteristics. Notably, there is a greater difference in how each model represents wind generation than PV, meaning the balanced scenario reduces these wind-related discrepancies, bringing the datasets closer together.

Across Fig.\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets. For instance, in the current scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets. This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influences drought duration estimates. Additionally, in all four panels, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=0.9\textwidth]{droughts_return_periods}
	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it. The horizontal dashed line in the two bottom panels marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset.}
	\label{fig:return_periods}
\end{figure}

@@

Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the wind panel (Fig.\ref{fig:res_droughts_seasonality}a). However, in this panel, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario. This reduction can be explained by the contribution of PV generation during the summer months in the existing scenario, even though it constitutes only 11\% of total capacity. Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours. In the projected scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency. The balanced mix of wind and PV in this scenario reduces the seasonal signal, slightly increasing the percentage of RES drought hours in winter months while significantly decreasing it in summer. Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N. \textcolor{magenta}{This is a good point to mention the seasonality of the demand in Ireland that Damian brought up, although it could also go in the conclusions}

@@

\textcolor{magenta}{So, here the panel names are different as in the other ones, with existing and balanced. this is what I meant by consistency. Also, note the panel a in the figure is labelled as c. that should be changed. }

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{drouhgts_varying_threshold}
	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~current and d)~projected installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment. The y-axis indicates the annual number of events. Each line correspond to the median value and the first and third quartiles are shown in shaded area}
	\label{fig:number_days_threshold}
\end{figure}

@@

\begin{itemize}
	\item Remind differences between three datasets
	\begin{itemize}
		\item Atlite and C3S-E G: incorporate locations of RES farms
		\item Atlite models better follows EirGrid data (for Wind)
	\end{itemize}
	\item Point Arguments with examples from the figures
	\begin{itemize}
		\item The two C3S-E datasets show similar results
		\item Atlite stands out of these two datasets
	\end{itemize}
	\item Conclusion: Location of RES farms has less impact than the accuracy of modelling the CF!
	\\
	\item Impact of the 2030 scenario (Anti correlation between wind and PV) | Seasonality
	\\
	\item Risk for policymakers. The more accurate Atlite model produces more RES drought events and longer RES drought durations, an important consideration for risk-adverse decision making.
	\\
	\item Where to go next ?
	\begin{itemize}
		\item Looking at future climate data (Combine Climate and Energy scenarios)
		\item Studying broader region (Europe), which is more relevant as everything will be interconnected but more difficult as we will need to repeat verification for all EU countries
	\end{itemize}
\end{itemize}

This study has highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts. One of the most significant differences is how each dataset incorporates the specific locations of RES farms. Both Atlite and C3S-E G take into account the exact locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation. However, our analysis shows that the Atlite dataset generally performs better, particularly in its wind generation estimates, where it closely follows the EirGrid data. This suggests that while the inclusion of RES farm locations is important, the accuracy of the CF modeling \textcolor{magenta}{I would say power curve model instead of CF modelling} used by Atlite may be the determining factor in its superior performance.\textcolor{magenta}{Superior how? Is finding more droughts superior or is it reproducing reality? I reckon it's the second, but make it explicit. Also, there is no explicit mention of PV, if you mention wind explicitly (which I agree on doing) I would add PV. Even though it is replacing that "However" with "Even though this is true for PV, our analysis shows that for wind..."}

When analysing the results, we found that the two C3S-E datasets (C3S-E G and C3S-E N) consistently show similar outcomes across various figures \textcolor{magenta}{I would not refer to them as figures, but as analysis or something like that. But if you do choose the word analysis, do not start the sentence with "When analysing the results..." and instead use something like "The results show..."} (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods},~\ref{fig:res_droughts_seasonality}), suggesting that their differences in methodology have minimal impact on the results. In contrast, Atlite stands out in its ability to model wind generation, as evidenced by the more accurate correlation with EirGrid data seen in the verification process (Section \ref{sec:wind_verification}). This distinction is further underscored in the figures \textcolor{magenta}{Same as before, avoid referring to the figures. We do not discuss the result (that was done in results), we discuss what it implies}, where Atlite often reports higher or more accurate return periods and a greater number of RES droughts (Fig.~\ref{fig:boxplot_number_events},~\ref{fig:return_periods}), particularly in scenarios with increased PV capacity. \textcolor{magenta}{For me, the point of this paragraph is quite clear but does not really come across at the moment: Atlite is the best one at reproducing reality (as established in the previous paragraph). At the same time, a long-term analysis shows more droughts for Atlite than any other model. This means that model selection is key to correctly quantify the risk and avoid undersizing the reserve capacity.}

One of the most critical conclusions of this study is that the specific locations of RES farms, while important, have a lower impact on the accuracy of RES drought modelling than the precision of the CF models themselves. \textcolor{magenta}{Same as before, I would not refer to CF models, but instead wind power curve and PV panel models} Atlite's superior performance suggests that refining these models could lead to better forecasting and planning for RES droughts, which is crucial for energy security.

Looking to the balanced scenario, our analysis suggests a significant shift in the dynamics of RES droughts due to the anti-correlation between wind and PV generation. This scenario reveals the importance of seasonality in RES droughts (Fig.~\ref{fig:res_droughts_seasonality}), where a diversified energy portfolio that includes both wind and PV could decrease the frequency and duration of RES droughts.
\textcolor{magenta}{I would merge these two paragraphs into one. Here are my thoughts in order: 1. We have found how wind and PV together perform better than individually in terms of droughts. 2. A specific mention of seasonality and how it is reduced. 3. Mention how Ireland has a very wind-dependent system and how targets for PV are ambitious in the coming years, almost making wind and PV capacities equal (maybe not the phrasing but you see the point, putting forward the balance again). 4. Link to policymakers in terms of how meeting targets of installed capacity can help improve the energy security thanks to this diversification. 5. Close with the importance of the right model selection for RES drought assessment, especially as more renewable capacity comes into the system.}
For policymakers, the findings of this study highlight the risks of inadequate modelling and planning. Accurate RES drought forecasting is essential for ensuring energy security, especially as Ireland moves toward an ambitious renewable energy target.

>>>>

\textcolor{magenta}{Some things to keep in mind when reading from Damian's comments:
	1. The whole current/projected and existing/balanced thing. Pick one and stick to it, including the figures (also fix the a/c level in fig 10)
	2. See if you are happy with the section titles throughout the document}

@@

This study aims to quantify the occurrence of RES drought events in Ireland for two energy scenarios. The first represents the existing energy mix, predominantly driven by wind, while the second is based on a projected energy scenarios, featuring a more balanced share between wind and PV. A uniform methodology is used to identify RES droughts for each of the three datasets, and differences in the frequency, the duration and the seasonality are presented.

@@

This study uses publicly available datasets to construct and validate the models for estimating the CF of wind and PV energy. The primary data sources include EirGrid, the transmission system operator for the ic of Ireland, and the ERA5 reanalysis dataset.

@@

EirGrid, which operates in the Republic of Ireland, and SONI, the system operator for Northern Ireland, provide detailed datasets on all wind and PV farms across the island of Ireland (Republic of Ireland and Northern Ireland) from 1990 to the present~\cite{eirgrid}. These datasets include critical information such as each farm’s installed capacity, name, and connection date. To enhance the accuracy of this data, the longitude and latitude for each farm were manually determined through online searches. For simplicity, this data will be referred to as originating from EirGrid, and the combined regions of the Republic of Ireland and Northern Ireland will be referred to as Ireland throughout the remainder of this document.

@@

This study investigates wind and PV capacity factors (CF) using three different models, resulting in six individual CF time series based on the installed capacities of 2023. In addition to analysing wind and PV generation separately, a combined CF is computed for each model by averaging wind and PV generation, weighted by their installed capacities at the end of 2023 (6.7 GW for wind and 0.6 GW for PV). This results in a total of nine CF time series for the existing scenario, which represents the existing distribution of RES installations in Ireland.

Given that PV capacity in Ireland remains relatively low in 2023, and to explore how a more balanced distribution of wind and PV might impact RES droughts, this study also considers a balanced scenario. In this scenario, the installed PV capacity is assumed to increase to 8 GW, while wind capacity rises to 9 GW, based on targets outlined in the roadmap published by EirGrid and SONI~\cite{eirgrid2023future}. For PV, new time series are generated for both the Atlite and C3S-E G models, incorporating a revised distribution of installed capacity across Ireland as specified in the roadmap. For wind, the CF time series remain unchanged, as significant shifts in wind farm distribution are not expected. The PV and wind time series are then combined using updated weights that reflect the installed capacities in the balanced scenario (9 GW for wind and 8 GW for PV).

@@

\begin{figure}[ht!]
	\centering
	\includegraphics{droughts_methodology}
	\caption{Wind time series of CF (red) and its 24-hour moving average (green) from the 7th to the 15th of July 2021. The black dashed line indicates the CF threshold. The grey bar shows the period identified as a wind drought under our definition}
	\label{fig:find_res_droughts}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{verification_power_curve}
	\caption{a/ Power curves of the Enercon E112.4500 with a 0.3 smoothing filter used by Atlite (blue) and the Vestas V136/3450 used by C3S-E (orange) b/ Histograms of wind CF for Ireland from Atlite (blue), C3S-E (orange) and EirGrid (shaded)}
	\label{fig:power_curve}
\end{figure}

The smoothing of the wind turbine power curve represents losses associated with each individual turbine and the interaction between the different turbines, which is important when modelling wind energy on larger scales. The histogram in Fig.~\ref{fig:power_curve}b shows that the C3S-E power curve tends to underestimate low CF values and overestimate higher ones, whereas the smoothed Atlite power curve more closely follows the recorded wind availability data from EirGrid.

@@

@@

The median, first, and third quartiles for the Atlite dataset are consistently greater than or equal to those of the other two datasets, regardless of the duration range or type of renewable energy considered. This difference arises from the wind turbine power curve model used in the C3S-E datasets, which tends to overestimate the wind CF (Fig.~\ref{fig:wind_verification_contour}). As a result, the overall number of RES droughts is underestimated in the C3S-E datasets compared to Atlite.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_number_events}
	\caption{Annual number of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents duration ranges in days (lower bound included), while the y-axis indicates the annual number of events. The boxes display the first and third quartiles and the median is marked by a black line. The whiskers indicate the 5th and 95th percentiles}
	\label{fig:boxplot_number_events}	
\end{figure}

@@

For PV (Fig.~\ref{fig:return_periods}b), Atlite behaves significantly different than the two C3S-E datasets. The Atlite results follow the ones for wind with a log-linear increase, but reach higher values in general with the longest event lasting forty days. For C3S-E G and C3S-E N, the duration of RES droughts increases in a log-linear pattern for events lasting less than 16 days. Beyond this duration, there is a sharp rise in drought duration for events up to a one-year return period. This sudden increase reflects the impact of winter on PV generation in Ireland, as PV output often remains below the CF threshold for extended periods during winter months. The difference between Atlite and the C3S-E datasets likely arises from how each model handles PV output near the threshold. Atlite appears to hover slightly above the threshold more frequently during these conditions, leading to shorter, more fragmented drought events. In contrast, C3S-E G and C3S-E N tend to fall below the threshold in similar conditions, resulting in longer continuous drought periods, especially during winter. This sensitivity to the threshold highlights how slight model differences can have substantial effects on drought duration estimates, particularly for PV in low-generation conditions.

For the existing installed capacity scenario (Fig.\ref{fig:return_periods}c), the return periods mirror those of wind (Fig.\ref{fig:return_periods}a), reinforcing that wind is the primary driver of drought patterns in the existing energy mix. In the balanced scenario (Fig.~\ref{fig:return_periods}d), the return periods for RES droughts increase across all durations. For example, the return period for a five-day drought event extends from roughly six months to four years in the Atlite dataset, and from about fifteen months to around five years in the two C3S-E datasets. The return periods for Atlite align more closely with those of C3S-E in the balanced scenario compared to the existing one. This convergence is due to the reduced proportion of wind capacity in the balanced scenario, which lessens the impact of wind-specific characteristics. Notably, there is a greater difference in how each model represents wind generation than PV, meaning the balanced scenario reduces these wind-related discrepancies, bringing the datasets closer together.

Across Fig.\ref{fig:return_periods}a, c, and, d, the return periods in the Atlite dataset are consistently higher than those in the two C3S-E datasets. For instance, in the existing scenario (Fig.\ref{fig:return_periods}c), an event with a one-year return period lasts six days in the Atlite dataset, compared to only five days in the C3S-E datasets. This difference underscores the importance of model selection when quantifying RES droughts, as each model’s assumptions and parametrisations significantly influences drought duration estimates. Additionally, in all four panels, the similarity between results from the two C3S-E datasets suggests that assumptions in the Atlite model—such as wind turbine power curve selection and PV panel specifications—have a greater impact on RES drought duration estimates than the precise geographic distribution of RES farms when studying the return periods of RES droughts.

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_return_periods}
	\caption{Return periods of the duration of RES droughts for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represents the return period time in a log-scale and the y-axis indicates the duration of RES drought associated with it. The horizontal dashed line in the two bottom panels marks the 5-day return period, with coloured vertical dashed marking its return period for each dataset.}
	\label{fig:return_periods}
\end{figure}

@@

Similar to previous results, the existing installed capacity wind-PV combination (Fig.\ref{fig:res_droughts_seasonality}c) shows patterns comparable to the wind panel (Fig.\ref{fig:res_droughts_seasonality}a). However, in this panel, the number of hours classified as RES droughts in summer decreases slightly compared to the wind-only scenario. This reduction can be explained by the contribution of PV generation during the summer months in the existing scenario, even though it constitutes only 11\% of total capacity. Since the number of RES drought hours for PV in summer is near zero, this small contribution has a noticeable impact on reducing overall drought hours. In the balanced scenario (Fig.~\ref{fig:res_droughts_seasonality}d), all three datasets show a reduction in monthly RES drought frequency. Annual reductions in median RES drought frequency are observed across the datasets, dropping from 14\% to 5\% for Atlite, from 8\% to 3\% for C3S-E G, and from 9\% to 4\% for C3S-E N. The balanced mix of wind and PV in this scenario reduces the seasonal signal, slightly increasing the percentage of RES drought hours in winter months while significantly decreasing it in summer. This shift is especially relevant, as energy demand is generally higher in winter than in summer, making winter droughts more critical in terms of grid security. By reducing the frequency of summer droughts while maintaining resilience during winter, the balanced scenario offers a more reliable energy supply that better aligns with seasonal demand patterns.

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=\textwidth]{droughts_varying_threshold}
	\caption{Number of RES drought per year for a)~Wind, b)~PV, and the combination for the c)~existing and d)~balanced installed capacity, for Atlite (red), C3S-E G (blue), and C3S-E N (purple). The x-axis represent the threshold from 0.05 to 0.2 of the CF with a 0.01 increment. The y-axis indicates the annual number of events. Each line correspond to the median value and the first and third quartiles are shown in shaded area}
	\label{fig:number_days_threshold}
\end{figure}

@@

This study has highlighted the key differences between the three datasets: Atlite, C3S-E G, and C3S-E N in their ability to model RES droughts. One of the most significant differences is how each dataset incorporates the specific locations of RES farms. Both Atlite and C3S-E G consider the locations of wind and PV farms, which should, in theory, provide a more accurate representation of RES generation. While this approach improves PV drought models, our analysis indicates that for wind, the Atlite dataset performs better overall, especially in its close alignment with EirGrid data for wind generation estimates. This finding suggests that although the inclusion of RES farm locations is beneficial, particularly for PV, the accuracy of the power curve model in Atlite may be the critical factor in producing realistic time series for low-generation events.

The results show that the two C3S-E datasets (C3S-E G and C3S-E N) consistently yield similar outcomes, indicating that their methodological differences have minimal impact. In contrast, Atlite stands out for its closer alignment with EirGrid data, particularly in wind generation, as seen in the verification process (Section \ref{sec:wind_verification}). This distinction is also evident in the analysis, where Atlite reports higher return periods and a greater number of RES droughts, especially in scenarios with balanced share of RES. One of the study's most critical findings is that the accuracy of RES drought modeling relies more on the precision of the wind power curve and PV panel models than on the specific locations of RES farms. Atlite’s superior performance highlights the importance of selecting accurate models for assessing RES drought risks. This careful model selection can better quantify risks, support effective planning, and avoid the potential underestimation of capacity needs, which is essential for ensuring energy security.

Looking at the balanced scenario, the analysis indicates a significant improvement in the management of RES droughts due to the complementary nature of wind and PV generation. Wind and PV together perform better in terms of drought frequency and duration than either would individually, largely because of the seasonal anti-correlation between the two energy sources. This diversification reduces the seasonal impact on RES droughts, as PV generation peaks in the summer and wind generation is more consistent in winter. Ireland currently has a highly wind-dependent energy system, but with ambitious targets for PV installations in the coming years, the energy mix is expected to approach a balance between wind and PV capacity. This balanced approach offers a more stable and secure energy supply by mitigating RES drought risks. For policymakers, these findings underscore the importance of meeting these capacity targets to enhance energy security through diversification. Additionally, the choice of model for RES drought assessment becomes increasingly critical as more renewable capacity is integrated into the system. Accurate forecasting and careful modelling are essential to support reliable energy planning in Ireland’s renewable future.
